The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}

### 2017-06-09 GRB111117A.tex
Address Andrews comments
Closes #7

We use the spectroscopic redshift, as well as new imaging data to constrain the
nature of the host galaxy as well as physical parameters of the GRB. The rest-frame X-ray derived hydrogen column density, for example, is high compared to a
complete sample of sGRBs and seems to follow the evolution with redshift as
traced by the hosts of long GRBs (lGRBs). This is consistent with the proportion
of the sGRB population originating in late-type galaxies, live in similar
environments to that of lGRB hosts.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs2.pdf}
	\caption{Imaging of the field of GRB~111117A with the X-shooter slit overlaid. Only one slit is shown, despite four epochs of spectroscopic observations because of the similarity in position angle. The image is the FORS2 $R$-band image for which the photometry is shown in Fig.~\ref{fig:SED}. 
	The blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. The insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
	\label{fig:spec_setup}
\end{figure}

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. 
%The X-ray afterglow light curve can be fit with a single power-law decay with an
%index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
%analysis using data obtained in photon counting (PC) mode in the widest time epoch where
%the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
%(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
%data) to prevent spectral changes that can affect the X-ray column density
%determination. 
The obtained spectrum is well described by an absorbed power-law
model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.

@@

Because the projected distance does not change much between $z = 1.3$ and $z =
2.211$, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013}
relating to host offset are unaffected.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is securely 
classified as belonging to the short class of GRBs. Because the observed classification
indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
\citep{Littlejohns2013a}, the updated redshift does not change this designation
significantly. 
%The intrinsic luminosity increases, as visible in the X-ray light curve
%in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
%the majority of long GRBs. 
\citet{Bromberg2013} investigated the degree to which the
long and short population distributions overlap and quantified the certainty in
the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
being a sGRB. Compared to the other two sGRB candidates at high redshift, GRB~060121
\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
($17_{-15}^{+14}$ per cent) and GRB~090426 \citep{Antonelli2009, Levesque2010,
	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
membership for GRB~111117A is much higher.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/NH_z.pdf}
	\caption{Rest frame, X-ray derived hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations. The detections are replaced with contours for clarity and the limits are shown with arrows.}
	\label{fig:NH_z}
\end{figure}

@@

%If we additionally assume that the hosts missing redshift in
%\citet{Berger2014} follow the redshift distribution of the hosts \textit{with}
%redshift, this implies that we are \textit{not} missing a dominant fraction of
%sGRB redshift at $z \approx 2$, due to host faintness.

@@

%In terms of Fig. 9 in \citet{Fong2013a}, the host is indeed  brighter than the average galaxies using the evolving luminosity function of galaxies. \todo{What is the average brightness of a galaxy at z = 2?} 
%\todo{Reformulate paragraph}

@@

@@

Using the new distance, the X-ray derived $n_H$ towards GRB~111117A is the
highest within a complete sample of sGRB hosts and is consistent with the
$n_H-z$ evolution traced by the hosts of lGRBs.The SFR of the host is in the
upper end of the sGRB host SFR distribution and despite the moderate age and
high $N_\mathrm{H}$, almost no dust is present.

>>>>

We use the spectroscopic redshift, as well as new imaging data to constrain the
nature of the host galaxy and the physical parameters of the GRB. The rest-frame X-ray derived hydrogen column density, for example, is high compared to a
complete sample of sGRBs and seems to follow the evolution with redshift as
traced by the hosts of long GRBs (lGRBs). This is consistent with a proportion
of the sGRB population originating in late-type galaxies, living in similar
environments to that of lGRB hosts.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs4.pdf}
	\caption{Imaging of the field of GRB~111117A with the X-shooter slit overlaid. Only one slit is shown, despite four epochs of spectroscopic observations because of the similarity in position angle. The image is the FORS2 $R$-band image for which the photometry is shown in Fig.~\ref{fig:SED}. 
	The blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. The insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
	\label{fig:spec_setup}
\end{figure}

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. 
%The X-ray afterglow light curve can be fit with a single power-law decay with an
%index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
%analysis using data obtained in photon counting (PC) mode in the widest time epoch where
%the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
%(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
%data) to prevent spectral changes that can affect the X-ray column density
%determination. 
The obtained spectrum is well described by an absorbed power-law
model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic equivalent hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a solar abundance and a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.

@@

Because the projected distance does not change significantly between $z = 1.3$ and $z =
2.211$, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013}
relating to host offset are unaffected.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is
securely classified as belonging to the short class of GRBs. Because the
observed classification indicators, $T_{90}$ and hardness ratio, do not depend
strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not
change this designation. %The intrinsic luminosity increases, as
% visible in the X-ray light curve
%in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
%the majority of long GRBs.
\citet{Bromberg2013} investigated the degree to which the long and short
population distributions overlap and quantified the certainty in the class
membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
percent probability of being a sGRB. Compared to the other two sGRB candidates
at high redshift, GRB~060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7
\lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB~090426
\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
per cent), the certainty in class membership for GRB~111117A is much higher.
This means that GRB~111117A is likely older by 1.8 Gyr than any other
non-collapsar GRB ever detected.

@@

For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
\citep{Lyman2017}, which is a challenge to reconcile  with the relatively large
offset from the host center derived in \citet{Margutti2012} and
\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
burst position that we don’t see. Along with the absence of dust, the large
offset from the host center indicates that the high $N_H$ arises because the
density in the GRB surrounding is high, or because the light from the afterglow
traverses a region of star formation.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/NH_z.pdf}
	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations. The detections are replaced with contours for clarity and the limits are shown with arrows.}
	\label{fig:NH_z}
\end{figure}

@@

@@

@@

In practice most NS-NS binaries will be eccentric at formation because of the SN
natal kicks, and this can be quite a big effect on merger times, although
eccentric systems normally circularise first. It could also be a BH-NS system,
and as the merger time scales roughly as $M^3$ the mass here almost as big an
effect as the separation ($a^4$).

@@

Using the new distance, the X-ray derived $n_H$ towards GRB~111117A is the
highest within a complete sample of sGRB hosts and is consistent with the
$n_H-z$ evolution traced by the hosts of lGRBs.The SFR of the host is in the
upper end of the sGRB host SFR distribution and despite the moderate age and
high $N_\mathrm{H}$, almost no dust is present. The high $N_H$ is difficult to
reconcile with the large projected host offset and the absence of dust. One
possible explanation could be, that GRB~111117A is formed through the prompt
channel of sGRBs \todo{ref} and originates in a star forming region located in
the outskirts of the host.
